Simulating DeLosh 1997

### 2023-06-15 Simulations/DeLosh97_Sim.qmd
sidebar updates2

::: {.column-margin}
###### Input Activation
$$
a_i(X)=\exp \left|-\gamma \cdot\left[X-X_i\right]^2\right|
$$

###### Output activation
$$
o_j(X)=\Sigma_{i=1, M} w_{j i} \cdot a_i(X) 
$$

###### Output Probability
$$
P\left[Y_j \mid X\right]=o_j(X) / \Sigma_{k=1, L} o_k(X) 
$$

###### Mean Response
$$
m(X)=\Sigma_{j=1, L} Y_j \cdot P\left[Y_j \mid X\right] 
$$

:::

### Generate Response
```{r}
#| code-fold: show
#| code-summary: "Toggle Code"
alm.response <- function(input=1,c) {
input.activation <- exp(-c*(input.layer - input)^2)
input.activation <<- input.activation/sum(input.activation)
#print(length(input.activation)); print(dim(weight.mat))
output.activation <<- weight.mat %*% input.activation
output.probability <<- output.activation/sum(output.activation)
mean.response <<- sum(output.layer * output.probability)
mean.response
}
```

\ \ \ \ \

### Update Weights Based on Feedback
```{r}
#| code-fold: show
#| code-summary: "Toggle Code"
#| code-overflow: wrap
alm.update <- function(corResp,c,lr){
  fz <- exp(-c*(output.layer - corResp)^2)
  teacherSignal <- (fz - output.activation)*lr
  #print(length(teacherSignal)); print(length(fz))
  wChange <- teacherSignal %*% t(input.activation)
  weight.mat <<- weight.mat + (wChange)
  weight.mat[weight.mat<0]=0 # prevent negative values
 # weight.mat[weight.mat>1]=1
  weight.mat <<- weight.mat
}

@@

::: {.column-margin}
###### Feedback Signal

@@

###### Weight Updates
$$
w_{ji}(t+1)=w_{ji}(t)+\alpha \cdot {f_i(Z(t))-O_j(X(t))} \cdot a_i(X(t))
$$
:::

### Exam Generalization
```{r}
#| code-fold: show
#| code-summary: "Toggle Code"
exam.response <- function(input,c){
  # Find the index of the input node with the highest activation
  trainVec = sort(unique(xt))
  nearestTrain <- trainVec[which.min(abs(input - trainVec))]
  aresp <- alm.response(nearestTrain,c)
  #max.index <- which.max(input.activation)
  xUnder = ifelse(min(trainVec) == nearestTrain, nearestTrain, trainVec[which(trainVec == nearestTrain) - 1])
  xOver = ifelse(max(trainVec) == nearestTrain, nearestTrain, trainVec[which(trainVec == nearestTrain) + 1])

@@

::: {.column-margin}
###### Input node actvation
$$
P[X_i|X] = \frac{a_i(X)}{\\sum_{k=1}^Ma_k(X)}
$$

###### Slope Computation
$$
E[Y|X_i]=m(X_i) + \bigg[\frac{m(X_{i+1})-m(X_{i-1})}{X_{i+1} - X_{i-1}} \bigg]\cdot[X-X_i]
$$
:::

@@

### Simulate Training:
```{r fig.width=10, fig.height=8}
#| column: page-inset-right
# split by type, then send each training dataset to simulation function
lowSim <- map(envTypes, ~ alm.sim(lowTrain %>% filter(type == .x), c = 1.4, lr = .4))

@@

Predictions for Generalization
```{r fig.width=11,fig.height=10}
#| column: page-inset-right
#| 
lowSimTest <- map_dfr(lowSim,simOrganize) %>% mutate(density = "low")
medSimTest <- map_dfr(medSim,simOrganize) %>% mutate(density = "med")
highSimTest <- map_dfr(highSim,simOrganize) %>% mutate(density = "high")

@@

Collpasing Across Density Levels gives us: 
```{r}
#| eval: false
#| column: page-inset-right
simTestAll %>% group_by(type,model,x,y) %>% summarise(resp=mean(resp))  %>% ggplot(aes(x=x,y=y)) + 
  geom_point(aes(x=x,y=resp,shape=model,color=model),alpha=.7,size=1) + 
  geom_line(aes(x=x,y=y),alpha=.4)+ 
  facet_grid(~type) + 
  theme_bw() + theme(legend.position="bottom")
```

@@

### Simulations with noise
```{r fig.height=10,fig.width=11}
#| eval: false
k = full.sim(c=1.4,lr=.4,noise=2.0)
k4 = full.sim(c=1.4,lr=.4,noise=4.0)

>>>>

::: column-margin
###### Input Activation

$$
a_i(X)=\exp \left|-\gamma \cdot\left[X-X_i\right]^2\right|
$$

###### Output activation

$$
o_j(X)=\Sigma_{i=1, M} w_{j i} \cdot a_i(X) 
$$

###### Output Probability

$$
P\left[Y_j \mid X\right]=o_j(X) / \Sigma_{k=1, L} o_k(X) 
$$

###### Mean Response

$$
m(X)=\Sigma_{j=1, L} Y_j \cdot P\left[Y_j \mid X\right] 
$$
:::

### Generate Response

```{r}
#| code-fold: show
#| code-summary: "Toggle Code"
alm.response <- function(input=1,c) {
input.activation <- exp(-c*(input.layer - input)^2)
input.activation <<- input.activation/sum(input.activation)
#print(length(input.activation)); print(dim(weight.mat))
output.activation <<- weight.mat %*% input.activation
output.probability <<- output.activation/sum(output.activation)
mean.response <<- sum(output.layer * output.probability)
mean.response
}
```

### Update Weights Based on Feedback

```{r}
#| code-fold: show
#| code-summary: "Toggle Code"
#| code-overflow: wrap
alm.update <- function(corResp,c,lr){
  fz <- exp(-c*(output.layer - corResp)^2)
  teacherSignal <- (fz - output.activation)*lr
  #print(length(teacherSignal)); print(length(fz))
  wChange <- teacherSignal %*% t(input.activation)
  weight.mat <<- weight.mat + (wChange)
  weight.mat[weight.mat<0]=0 # prevent negative values
 # weight.mat[weight.mat>1]=1
  weight.mat <<- weight.mat
}

@@

::: column-margin
###### Feedback Signal

@@

###### Weight Updates

$$
w_{ji}(t+1)=w_{ji}(t)+\alpha \cdot {f_i(Z(t))-O_j(X(t))} \cdot a_i(X(t))
$$
:::

### Exam Generalization

```{r}
#| code-fold: show
#| code-summary: "Toggle Code"
exam.response <- function(input,c){
  # Find the index of the input node with the highest activation
  trainVec = sort(unique(xt))
  nearestTrain <- trainVec[which.min(abs(input - trainVec))]
  aresp <- alm.response(nearestTrain,c)
  #max.index <- which.max(input.activation)
  xUnder = ifelse(min(trainVec) == nearestTrain, nearestTrain, trainVec[which(trainVec == nearestTrain) - 1])
  xOver = ifelse(max(trainVec) == nearestTrain, nearestTrain, trainVec[which(trainVec == nearestTrain) + 1])

@@

::: column-margin
###### Input node actvation

$$
P[X_i|X] = \frac{a_i(X)}{\\sum_{k=1}^Ma_k(X)}
$$

###### Slope Computation

$$
E[Y|X_i]=m(X_i) + \bigg[\frac{m(X_{i+1})-m(X_{i-1})}{X_{i+1} - X_{i-1}} \bigg]\cdot[X-X_i]
$$
:::

@@

### Simulate Training:

```{r fig.width=10, fig.height=8}
#| column: page-inset-right
# split by type, then send each training dataset to simulation function
lowSim <- map(envTypes, ~ alm.sim(lowTrain %>% filter(type == .x), c = 1.4, lr = .4))

@@

Predictions for Generalization

```{r fig.width=11,fig.height=10}
#| column: page-inset-right
#| 
lowSimTest <- map_dfr(lowSim,simOrganize) %>% mutate(density = "low")
medSimTest <- map_dfr(medSim,simOrganize) %>% mutate(density = "med")
highSimTest <- map_dfr(highSim,simOrganize) %>% mutate(density = "high")

@@

Collpasing Across Density Levels gives us:

```{r}
#| eval: false
#| column: page-inset-right
simTestAll %>% group_by(type,model,x,y) %>% summarise(resp=mean(resp))  %>% ggplot(aes(x=x,y=y)) + 
  geom_point(aes(x=x,y=resp,shape=model,color=model),alpha=.7,size=1) + 
  geom_line(aes(x=x,y=y),alpha=.4)+ 
  facet_grid(~type) + 
  theme_bw() + theme(legend.position="bottom")
```

@@

### Simulations with noise

```{r fig.height=10,fig.width=11}
#| eval: false
k = full.sim(c=1.4,lr=.4,noise=2.0)
k4 = full.sim(c=1.4,lr=.4,noise=4.0)
